---
title: 'SNAG: Correlations'
author: "Shuying Yu"
date: "June 18, 2020"
output: pdf_document
---

```{r, include=FALSE}
library(foreign)
#install.packages("compute.es")
library(compute.es)
#install.packages("Rcpp")
#install.packages("WRS2")
library(WRS2)
#install.packages("pastecs")
library(pastecs)
#install.packages("multcomp")
library(multcomp)
#install.packages("emmeans")
library(emmeans)
#install.packages("effects")
library(effects)
#install.packages("lsr")
library(lsr)
#install.packages("lsmeans")
library(lsmeans)
#install.packages("rlang")
#install.packages("tibble")
#install.packages("dplyr")
library(haven)
library(labelled)
library(psych)
library(Rmisc)
library(car)
library(tidyverse)
library(agricolae)
library(afex)
library(lm.beta)
library(rlang)
#install.packages("sjstats")
library(sjstats)
#install.packages("effsize")
library(effsize)
library(ggplot2)
library(ggpubr)
library(Hmisc)
library(rstatix)
#install.packages("ez")
library(ez)
library(corrplot)
library(PerformanceAnalytics)
```

#Variables

Because correlations require observations with the same vector length, this was done for participants who have all tasks completed.

**Dependent variables:**

Names of variables and brief description

* SBSOD (Santa Barbara Sense of Direction scale)

* SOT (estimated angle [SOT_Angle] and angular error [SOT_Error]; if anything, just SOT_Angle is the measure we care about)

* MRM (Money Road Map)

* DSP_Wayfinding (proportion correct)

* DSP_SI (solution index strict coding)

* MAZE_Wayfinding (proportion correct)

* MAZE_Moves (number of moves during exploration)

* LOOP_MeanPositionErr (averge meters across 3 radii)

* LOOP_PositionPSTD (average PSTD for position error across 3 radii)

* LOOP_MeanDegrees (average degrees across 3 radii)

* LOOP_DegreesPSTD (average PSTD for degrees traveled across 3 radii)


#Data

```{r}
#Load data
corr_dat <- read.csv("Correlations_ALL.csv")

#Remove SubjectID column
corr_dat <- corr_dat[,c(2:13)]

#Show first 10 rows/subjects
head(corr_dat, 10)
```

#Correlation matrix with p-values

```{r}
#Save correlation matrix into variable
res <- rcorr(as.matrix(corr_dat), type = "pearson")
res
```

#Table

Formatting a correlation matrix into a table with 4 columns containing :

* Column 1 : row names (variable 1 for the correlation test)

* Column 2 : column names (variable 2 for the correlation test)

* Column 3 : the correlation coefficients

* Column 4 : the *p*-values of the correlations


```{r}
#Create function to combine coefficients and p-values

#Code modified from Reference [1] and [2]

#coeff_mat: matrix of the correlation coefficients
#p_mat : matrix of the correlation p-values

CorMat <- function(coeff_mat, p_mat) {
  
  #Save upper half triangle of coefficients
  #This is because the other half is redundant
  up_half <- upper.tri(coeff_mat)
  
  #Save into dataframe
  data.frame(
    
    #Variable 1
    Var1 = rownames(coeff_mat)[row(coeff_mat)[up_half]],
    #Variable 2
    Var2 = rownames(coeff_mat)[col(coeff_mat)[up_half]],
    #Correlation coefficient
    Corr  =(coeff_mat)[up_half],
    #p-value
    p_value = p_mat[up_half]
    )
}
```


```{r}
#Pearson correlation
res2 <- rcorr(as.matrix(corr_dat), type="pearson")

#Use function to place into table

#res2$r = correlation coefficient
#res2$P = associated p-value
CorMat(res2$r, res2$P)
```


#Visualize correlation matrix

##Scatter and chart

```{r}
#Scatter plot, histogram, and correlation value
chart.Correlation(corr_dat, histogram=TRUE, pch=19)
```


In the above plot:

* The distribution of each variable is shown on the diagonal.

* On the bottom of the diagonal : the bivariate scatter plots with a fitted line are displayed

* On the top of the diagonal : the value of the correlation plus the significance level as stars

* Each significance level is associated to a symbol : *p*-values(0, 0.001, 0.01, 0.05, 0.1, 1) <=> symbols(`***`, `**`, `*`, `.`, ` `)

In the pointing task (SOT), angle and error are negatively correlated. The more accurate subjects are at selecting the correct able, the less errors they make.

From the distribtion of the historgram, most subjects tend to score high on the SOT angle and on the MRM. 

We can see that the wayfinding measures from DSP and MAZE are significantly positively correlated, indicating that performance in successfully and accurately completing trials are related in both tasks. 

Variability in position error and degrees traveled are also significantly correlated.


##Correlation

```{r}
#Display only "upper" triangular portion to cut redundancy

corrplot(res2$r, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
```

```{r}
#Show correlation coefficient numbers

corrplot(res2$r, method="number", number.cex=0.5)
```


```{r}
#Significance at p = 0.05

#Insignificant correlations are left blank
corrplot(res2$r, type="upper", order="hclust", 
         p.mat = res2$P, sig.level = 0.05, insig = "blank")
```


```{r}
#Numbers are correlation coefficients
#Colored squares are those that meet significance (alpha = 0.05)

#Color palette and code taken from Reference [1] and [2]
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))

corrplot(res2$r, method="color", col=col(200),  
         type="upper", order="hclust", 
         #Add coefficient of correlation
         addCoef.col = "black",
         #Text label color and rotation
         tl.col="black", tl.srt=45, 
         # Combine with significance
         p.mat = res2$P, sig.level = 0.05, insig = "blank", number.cex=0.6,
         #Hide correlation coefficient on the principal diagonal
         diag=FALSE 
         )
```


Positive correlations are displayed in blue and negative correlations in red color. Color intensity and the size of the circle are proportional to the correlation coefficients. On the right side of the correlogram, the legend color shows the correlation coefficients and the corresponding colors.

Overall, the DSP and MAZE wayfinding success measures are the most interesting (but expected) association. In general, there are no other major associations of interest or concern.


#References

1. Correlation matrix : A quick start guide to analyze, format and visualize a correlation matrix using R softwards. http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-analyze-format-and-visualize-a-correlation-matrix-using-r-software

2. Visualize correlation matrix using correlogram. http://www.sthda.com/english/wiki/visualize-correlation-matrix-using-correlogram







