---
title: "Yu et al Psych Science - Data Analysis"
author: "Shuying Yu"
date: "April 21, 2020"
output: pdf_document
---

#SAMPLE SIZES 

**TOTAL**

* Young = 85 
    + 39 female
    + 46 males

* Old = 66 
    + 40 females
    + 26 males


**LOOP:**

* Young = 52
    + 25 females
    + 27 males

* Old = 56
    + 33 females
    + 23 males

**MAZE:**

* Young = 50
    + 26 females
    + 24 males

* Old = 39
    + 25 females
    + 14 males


**DSP:**

* Young = 54
    + 26 females
    + 28 males

* Old = 40
    + 19 females
    + 21 males


```{r, include=FALSE}
library(foreign)
#install.packages("compute.es")
library(compute.es)
#install.packages("Rcpp")
#install.packages("WRS2")
library(WRS2)
#install.packages("pastecs")
library(pastecs)
#install.packages("multcomp")
library(multcomp)
#install.packages("emmeans")
library(emmeans)
#install.packages("effects")
library(effects)
#install.packages("lsr")
library(lsr)
#install.packages("lsmeans")
library(lsmeans)
#install.packages("rlang")
#install.packages("tibble")
#install.packages("dplyr")
library(haven)
library(labelled)
library(psych)
library(Rmisc)
library(car)
library(tidyverse)
library(agricolae)
library(afex)
library(lm.beta)
library(rlang)
#install.packages("sjstats")
library(sjstats)
#install.packages("effsize")
library(effsize)
library(ggplot2)
library(ggpubr)
library(Hmisc)
library(rstatix)
#install.packages("ez")
library(ez)
```

***

#LOOP


##Position Error

Length of the straight line distance between the actual start of the circle and where participants thought was the start of the circle.

Expressed in meters (m).

\hfill\break

```{r}
#Read dataframe
df_loop <- read.csv("LOOP_ALL_2020_removed.csv")
names(df_loop)[1] <- "SubjectID"
df_loop$SubjectID <- as.factor(df_loop$SubjectID)
```


```{r}
#Load the data and show some random ros by groups
set.seed(123)
df_loop %>% sample_n_by(Group, Sex, size=1)
```


```{r}
#Gather columns meandist1,2,3 into long format

#convert SubjectID into factor

df_loop <- df_loop %>% 
  gather(key="Radius", value="PE", mean_dist1, mean_dist2, mean_dist3) %>% 
  convert_as_factor(SubjectID, Radius)

#Inspect some random rows of the data by groups
set.seed(123)
df_loop %>% sample_n_by(Group, Sex, Radius, size=1)
```


```{r}
aov_pe <- anova_test(
  data=df_loop, dv=PE, wid=SubjectID,
  within=c(Radius),
  between=c(Group, Sex),
  effect.size = "pes"
  )

#get_anova_table(aov_pe)
aov_pe
```


```{r}
library(apaTables)
get.ci.partial.eta.squared(5.46, 1, 36, conf.level = .95)
```



###Post-hoc

#####Sex:Radius

```{r}
# Two-way ANOVA at each age group level
# Effect of Sex at each level of Radius
one.way <- df_loop %>%
  group_by(Radius) %>%
  anova_test(dv = PE, wid = SubjectID, between = Sex) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
one.way
```



#####Group:Radius

```{r}
# Two-way ANOVA at each age group level
# Effect of Group at each level of Radius
one.way2 <- df_loop %>%
  group_by(Radius) %>%
  anova_test(dv = PE, wid = SubjectID, between = Group) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
one.way2
```



##Position Error Variability

```{r}
#Read dataframe
df_loop1.1 <- read.csv("LOOP_ALL_2020_removed.csv")
names(df_loop1.1)[1] <- "SubjectID"
```


```{r}
#Load the data and show some random ros by groups
set.seed(123)
df_loop1.1 %>% sample_n_by(Group, Sex, size=1)
```


```{r}
#Gather columns meandist1,2,3 into long format

#convert SubjectID into factor

df_loop1.1 <- df_loop1.1 %>% 
  gather(key="Radius", value="PE_pstd", pstd_dist1, pstd_dist2, pstd_dist3) %>% 
  convert_as_factor(SubjectID, Radius)

#Inspect some random rows of the data by groups
set.seed(123)
df_loop1.1 %>% sample_n_by(Group, Sex, Radius, size=1)
```


```{r}
aov_pestd <- anova_test(
  data=df_loop1.1, dv=PE_pstd, wid=SubjectID,
  within=c(Radius),
  between=c(Group, Sex),
  effect.size = "pes"
  )

#get_anova_table(aov_pe)
aov_pestd
```

##Degrees Traveled

```{r}
#Read dataframe
df_loop2 <- read.csv("LOOP_ALL_2020_removed.csv")
names(df_loop2)[1] <- "SubjectID"
```


```{r}
#Load the data and show some random ros by groups
set.seed(123)
df_loop2 %>% sample_n_by(Group, Sex, size=1)
```



```{r}
#Gather columns meandegrees1,2,3 into long format

#convert SubjectID into factor


df_loop2 <- df_loop2 %>% 
  gather(key="Radius", value="DE", mean_degrees1, mean_degrees2, mean_degrees3) %>% 
  convert_as_factor(SubjectID, Radius)

#Inspect some random rows of the data by groups
set.seed(123)
df_loop2 %>% sample_n_by(Group, Sex, Radius, size=1)
```


```{r}
aov_de <- anova_test(
  data=df_loop2, dv=DE, wid=SubjectID,
  within=c(Radius),
  between=c(Group, Sex),
  effect.size = "pes"
  )

#get_anova_table(aov_de)
aov_de
```



##Degrees Traveled Variability


```{r}
#Read dataframe
df_loop2.1 <- read.csv("LOOP_ALL_2020_removed.csv")
names(df_loop2.1)[1] <- "SubjectID"
```


```{r}
#Load the data and show some random ros by groups
set.seed(123)
df_loop2.1 %>% sample_n_by(Group, Sex, size=1)
```


```{r}
#Gather columns meandegrees1,2,3 into long format

#convert SubjectID into factor


df_loop2.1 <- df_loop2.1 %>% 
  gather(key="Radius", value="DE_pstd", pstd_degrees1, pstd_degrees2, pstd_degrees3) %>% 
  convert_as_factor(SubjectID, Radius)

#Inspect some random rows of the data by groups
set.seed(123)
df_loop2.1 %>% sample_n_by(Group, Sex, Radius, size=1)
```



```{r}
aov_destd <- anova_test(
  data=df_loop2.1, dv=DE_pstd, wid=SubjectID,
  within=c(Radius),
  between=c(Group, Sex),
  effect.size = "pes"
  )

#get_anova_table(aov_de)
aov_destd
```



***


#MAZE

```{r}
df_maze <- read.csv("MAZE_ALL_2020.csv")
```


##Test against Chance Level

```{r}
#Young female
df_maze_yf <- df_maze %>% filter(Group == "Young" & Sex == "Female")
t.test(df_maze_yf$AvgAccuracy, mu=.1111, alternative="greater", conf.level=0.95)
```


```{r}
mean(df_maze_yf$AvgAccuracy)
sd(df_maze_yf$AvgAccuracy)
CI(df_maze_yf$AvgAccuracy, ci = 0.95)

#Cohen's d: (mean - mu)/sd
(mean(df_maze_yf$AvgAccuracy) - .1111)/sd(df_maze_yf$AvgAccuracy)

cohen.d.ci((mean(df_maze_yf$AvgAccuracy) - .1111)/sd(df_maze_yf$AvgAccuracy),
            n1=26, alpha = 0.05)
```


```{r}
#Young male
df_maze_ym <- df_maze %>% filter(Group == "Young" & Sex == "Male")
t.test(df_maze_ym$AvgAccuracy, mu=.1111, alternative="greater", conf.level=0.95)
```


```{r}
mean(df_maze_ym$AvgAccuracy)
sd(df_maze_ym$AvgAccuracy)
CI(df_maze_ym$AvgAccuracy, ci = 0.95)

#Cohen's d: (mean - mu)/sd
(mean(df_maze_ym$AvgAccuracy) - .1111)/sd(df_maze_ym$AvgAccuracy)

cohen.d.ci((mean(df_maze_ym$AvgAccuracy) - .1111)/sd(df_maze_ym$AvgAccuracy),
            n1=24, alpha = 0.05)
```




```{r}
#Midlife female
df_maze_mf <- df_maze %>% filter(Group == "Midlife" & Sex == "Female")
t.test(df_maze_mf$AvgAccuracy, mu=.1111, alternative="greater", conf.level=0.95)
```


```{r}
mean(df_maze_mf$AvgAccuracy)
sd(df_maze_mf$AvgAccuracy)
CI(df_maze_mf$AvgAccuracy, ci = 0.95)

#Cohen's d: (mean - mu)/sd
(mean(df_maze_mf$AvgAccuracy) - .1111)/sd(df_maze_mf$AvgAccuracy)

cohen.d.ci((mean(df_maze_mf$AvgAccuracy) - .1111)/sd(df_maze_mf$AvgAccuracy),
            n1=25, alpha = 0.05)
```


```{r}
#Midlife male
df_maze_mm <- df_maze %>% filter(Group == "Midlife" & Sex == "Male")
t.test(df_maze_mm$AvgAccuracy, mu=.1111, alternative="greater", conf.level=0.95)
```

```{r}
mean(df_maze_mm$AvgAccuracy)
sd(df_maze_mm$AvgAccuracy)
CI(df_maze_mm$AvgAccuracy, ci = 0.95)

#Cohen's d: (mean - mu)/sd
(mean(df_maze_mm$AvgAccuracy) - .1111)/sd(df_maze_mm$AvgAccuracy)

cohen.d.ci((mean(df_maze_mm$AvgAccuracy) - .1111)/sd(df_maze_mm$AvgAccuracy),
            n1=14, alpha = 0.05)
```


##Wayfinding Success / Accuracy


```{r}
options(contrasts=c("contr.sum", "contr.poly"))
interaction_acc <- aov(AvgAccuracy~Group*Sex, data = df_maze)
Anova(interaction_acc, type="III")
```


```{r}
acc <- Anova(interaction_acc, type="III")
anova_stats(acc)
```


```{r}
#Based on bootstrapping
eta_sq(acc, partial = TRUE, ci.lvl = .95,
       n = 1000, method = "quantile")
```



###Post-hoc

```{r}
# Effect of treatment at each level of Group
df_maze %>%
  group_by(Group) %>%
  anova_test(AvgAccuracy ~ Sex, 
             type = "III",
             effect.size = "pes") %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
```

```{r}
#Cohen's d

df_maze_midlife <- df_maze %>% filter(Group == "Midlife")
df_maze_young <- df_maze %>% filter(Group == "Young")
df_maze_female <- df_maze %>% filter(Sex == "Female")
df_maze_male <- df_maze %>% filter(Sex == "Male")

#Effect of Sex on Midlife
cohen.d(AvgAccuracy ~ Sex, group=Group, data=df_maze_midlife,
        conf.level=0.95, pooled=T, paired=F)
```


```{r}
#Effect of Sex on Young
cohen.d(AvgAccuracy ~ Sex, group=Group, data=df_maze_young,
        conf.level=0.95, pooled=T, paired=F)
```


```{r}
mean(df_maze_midlife$AvgAccuracy)
sd(df_maze_midlife$AvgAccuracy)
CI(df_maze_midlife$AvgAccuracy, ci = 0.95)

mean(df_maze_young$AvgAccuracy)
sd(df_maze_young$AvgAccuracy)
CI(df_maze_young$AvgAccuracy, ci = 0.95)
```


```{r}
# Effect of treatment at each level of Sex
df_maze %>%
  group_by(Sex) %>%
  anova_test(AvgAccuracy ~ Group,
             type = "III",
             effect.size = "pes") %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
```


```{r}
#Effect of Age on Female
cohen.d(AvgAccuracy ~ Group, group=Sex, data=df_maze_female,
        conf.level=0.95, pooled=T, paired=F)
```


```{r}
#Effect of Age on Male
cohen.d(AvgAccuracy ~ Group, group=Sex, data=df_maze_male,
        conf.level=0.95, pooled=T, paired=F)
```

```{r}
mean(df_maze_female$AvgAccuracy)
sd(df_maze_female$AvgAccuracy)
CI(df_maze_female$AvgAccuracy, ci = 0.95)

mean(df_maze_male$AvgAccuracy)
sd(df_maze_male$AvgAccuracy)
CI(df_maze_male$AvgAccuracy, ci = 0.95)
```


##Moves

###Correlation

```{r}
#95% CI here is for the estimate (correlation)
cor.test(df_maze$Moves, df_maze$AvgAccuracy, method="pearson")
```


```{r}
#needs alternative and conf level
cor.test(df_maze$Moves, df_maze$AvgAccuracy, method="pearson",
         alternative="two.sided", conf.level = .95)
```



###Mean and SD

```{r}
describe.by(df_maze$Moves)
```

```{r}
#95% CI for mean
CI(df_maze$Moves, ci = 0.95)
```


```{r}
#By Groups: Young vs Midlife
mean(df_maze_young$Moves)
sd(df_maze_young$Moves)
CI(df_maze_young$Moves, ci = 0.95)

mean(df_maze_midlife$Moves)
sd(df_maze_midlife$Moves)
CI(df_maze_midlife$Moves, ci = 0.95)

#By Sex: Female vs Male
mean(df_maze_female$Moves)
sd(df_maze_female$Moves)
CI(df_maze_female$Moves, ci = 0.95)

mean(df_maze_male$Moves)
sd(df_maze_male$Moves)
CI(df_maze_male$Moves, ci = 0.95)
```



###ANOVA

```{r}
options(contrasts=c("contr.sum", "contr.poly"))
interaction_moves <- aov(Moves~Group*Sex, data = df_maze)
Anova(interaction_moves, type="III")
```



```{r}
moves1 <- Anova(interaction_moves, type="III")
anova_stats(moves1)
```


```{r}
#Based on bootstrapping
eta_sq(moves1, partial = TRUE, ci.lvl = .95,
       n = 1000, method = "quantile")
```


##ANCOVA 

On Wayfinding Success/Accuracy controlling for Moves

```{r}
options(contrasts=c("contr.sum", "contr.poly"))
interaction_acc2 <- aov(AvgAccuracy~Moves + Group*Sex, data = df_maze)
Anova(interaction_acc2, type="III")
```


```{r}
acc2 <-Anova(interaction_acc2, type="III")
anova_stats(acc2)
```

```{r}
#Based on bootstrapping
eta_sq(acc2, partial = TRUE, ci.lvl = .95,
       n = 1000, method = "quantile")
```

###Post-hoc effect sizes with adjusted means

```{r}
#Adjusted means of wayfinding with associated 95%CI and SE
#Adjusted for the effect of the covariate Moves
adjustedmeans <- effect("Group*Sex", interaction_acc2, se=T)
summary(adjustedmeans)
adjustedmeans$se
```

SE order = midlife female, midlife male, young female, young male

```{r}
#Adjusted SD from adjusted means for Group*Sex
adjustedmeans$se*sqrt(c(25, 14, 26, 24))
```


```{r}
#Effect size and CI for Simple efect of Sex - Midlife (midlife women and men)

#Effect sizes for the difference between midlife women and men
mes( 0.1088111, 0.2686762, 0.2455819, 0.1699723, 25, 14)
```


```{r}
#Effect size and CI for Simple efect of Sex - Young (young women and men)

#Effect sizes for the difference between young women and men
mes( 0.2078208, 0.6539117, 0.3092134, 0.2343850, 26, 24)
```



```{r}
#Effect size and CI for Simple efect of age - women (young and midlife women)

#Effect sizes for the difference between midlife and yong women
mes(0.1088111, 0.2078208, 0.2455819, 0.3092134, 25, 26)
```


```{r}
#Effect size and CI for Simple efect of age - men (young and midlife men)

#Effect sizes for the difference between midlife and yong men
mes(0.2686762, 0.6539117, 0.1699723, 0.2343850, 14, 24)
```




###Post-hoc

```{r}
# Effect of treatment at each level of Group
df_maze %>%
  group_by(Group) %>%
  anova_test(AvgAccuracy ~ Moves + Sex,
             type = "III",
             effect.size = "pes") %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
```


```{r}
# Pairwise comparisons at level of Group
pwc.cov1 <- df_maze %>% 
  group_by(Group) %>%
  emmeans_test(
    AvgAccuracy ~ Sex, covariate = Moves,
    p.adjust.method = "bonferroni",
    detailed = TRUE
    )
pwc.cov1 
```



```{r}
# Effect of treatment at each level of Sex
df_maze %>%
  group_by(Sex) %>%
  anova_test(AvgAccuracy ~ Moves + Group,
             type="III",
             effect.size = "pes")  %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
```


```{r}
# Pairwise comparisons at level of Sex
pwc.cov2 <- df_maze %>% 
  group_by(Sex) %>%
  emmeans_test(
    AvgAccuracy_Percent ~ Group, covariate = Moves,
    p.adjust.method = "bonferroni"
    )
pwc.cov2 
```








***


#DSP


##Solution Index 

Propensity (proportion) or preference to take more shortcuts (range is 0-1); expressed in %

>More likely to take shortcuts = >0.50  

>Less likely to take shortcuts = <0.50

Formula: (# of STRICT shortcuts/# of successful trials)

\hfill\break


```{r}
#Read dataframe
df_dsp <- read.csv("DSP_ALL_2020_removed.csv")
```



```{r}
options(contrasts=c("contr.sum", "contr.poly"))
interaction_SI <- aov(SI_strict~Group*Sex, data = df_dsp)
Anova(interaction_SI, type="III")
```


```{r}
aov_si <- Anova(interaction_SI, type="III")
anova_stats(aov_si)
```

```{r}
#Based on bootstrapping
eta_sq(aov_si, partial = TRUE, ci.lvl = .95,
       n = 1000, method = "quantile")
```

###Post-hoc

```{r}
# Effect of treatment at each level of Group
df_dsp %>%
  group_by(Group) %>%
  anova_test(SI_strict ~ Sex, 
             type = "III",
             effect.size = "pes") %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
```

```{r}
df_dsp_yf <- df_dsp %>% filter(Group == "Young" & Sex == "Female")
df_dsp_ym <- df_dsp %>% filter(Group == "Young" & Sex == "Male")
df_dsp_mf <- df_dsp %>% filter(Group == "Midlife" & Sex == "Female")
df_dsp_mm <- df_dsp %>% filter(Group == "Midlife" & Sex == "Male")
df_dsp_female <- df_dsp %>% filter(Sex == "Female")
df_dsp_male <- df_dsp %>% filter(Sex == "Male")
df_dsp_young <- df_dsp %>% filter(Group == "Young")
df_dsp_midlife <- df_dsp %>% filter(Group == "Midlife")

mean(df_dsp_young$SI_strict)
sd(df_dsp_young$SI_strict)
CI(df_dsp_young$SI_strict, ci = 0.95)

mean(df_dsp_midlife$SI_strict)
sd(df_dsp_midlife$SI_strict)
CI(df_dsp_midlife$SI_strict, ci = 0.95)

mean(df_dsp_male$SI_strict)
sd(df_dsp_male$SI_strict)
CI(df_dsp_male$SI_strict, ci = 0.95)

mean(df_dsp_female$SI_strict)
sd(df_dsp_female$SI_strict)
CI(df_dsp_female$SI_strict, ci = 0.95)
```

```{r}
#Effect of Age on Female
cohen.d(SI_strict ~ Group, group=Sex, data=df_dsp_female,
        conf.level=0.95, pooled=T, paired=F)
```

```{r}
#Effect of Age on Male
cohen.d(SI_strict ~ Group, group=Sex, data=df_dsp_male,
        conf.level=0.95, pooled=T, paired=F)
```


```{r}
# Effect of treatment at each level of Sex
df_dsp %>%
  group_by(Sex) %>%
  anova_test(SI_strict ~ Group,
             type = "III",
             effect.size = "pes") %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
```

```{r}
#Effect of Sex on Young
cohen.d(SI_strict ~ Sex, group=Group, data=df_dsp_young,
        conf.level=0.95, pooled=T, paired=F)
```

```{r}
#Effect of Sex on Midlife
cohen.d(SI_strict ~ Sex, group=Group, data=df_dsp_midlife,
        conf.level=0.95, pooled=T, paired=F)
```




##Wayfinding Success / Proportion of successful trials

Number of trials they successed over total number of trials attempted which is usually 20 trials.

Expressed as a percent (range from 0-1).

```{r}
options(contrasts=c("contr.sum", "contr.poly"))
interaction_success <- aov(Success_dec~Group*Sex, data = df_dsp)
Anova(interaction_success, type="III")
```


```{r}
success1 <-  Anova(interaction_success, type="III")
anova_stats(success1)
```

```{r}
#Based on bootstrapping
eta_sq(success1, partial = TRUE, ci.lvl = .95,
       n = 1000, method = "quantile")
```


```{r}
#95%CI

mean(df_dsp_young$Success_dec)
sd(df_dsp_young$Success_dec)
CI(df_dsp_young$Success_dec, ci = 0.95)

mean(df_dsp_midlife$Success_dec)
sd(df_dsp_midlife$Success_dec)
CI(df_dsp_midlife$Success_dec, ci = 0.95)

mean(df_dsp_female$Success_dec)
sd(df_dsp_female$Success_dec)
CI(df_dsp_female$Success_dec, ci = 0.95)

mean(df_dsp_male$Success_dec)
sd(df_dsp_male$Success_dec)
CI(df_dsp_male$Success_dec, ci = 0.95)
```
























































































